---
title: "Using OpenCyto Command Line Tools"
author: "Greg Finak"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using OpenCytoCLI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Notes

Currently openCyto command line tools work with Linux and Mac OS X

## Install OpenCyto Command Line Tools

To install the package run the following in R:


```{r eval=FALSE,echo=TRUE}
library(devtools)
install_github("RGLab/opencytoCL")
```

You will need all the dependencies:

- openCyto
- optparse
- littler

## Linux and MacOS X

Then, in R:

```{r, eval=FALSE,results="asis"}
install_openCytoCL()
```

## Usage
For help, at the command prompt, type:

```{bash, comment = "",size="small",tidy=TRUE}
opencyto -h
```

### parse
Import a FlowJo workspace and related FCS files and create a GatingSet.

#### options

`<workspace>` the location of the workspace

`<fcs_read_from>` the location of the fcs files

`<save_to>` the name of the gating set to save

`--group=<group_number>` option to specify the group number to import. Print all group names using `load` (see below).

`--annotate=<keywords>` optionally annotate by importing keywords. Keywords should be passed in as a quoted, comma separated list. If parsing FCS files, the keywords are taken from the FCS files.  If parsing a workspace, the keywords are taken from the workspace. Spaces in keywords should be replaced with `:`.

`--stats` Extract population statistics and write them to a file.

### load
Load a FlowJo workspace and print the sample groups in therein.

#### options

`--gs=<gatingset>` the name of the gating set to load. Created using `parse` (see above) or `process` (see below) 

### gate
Gate a GatingSet using a template, saving the results. Optionally extract population statistics. Any existing gates in the GatingSet are removed. 

#### options

`--gs=<gatingset>`  A GatingSet on which to perform automated gating.

`--template=<template>`  The name of an openCyto template csv file.

`--stats`  Extract population statistics and write them to a file.

### process

`<fcs_read_from>` location of files to process

`<save_to>` name of gating set to save to

`--compensate` perform compensation using the matrix in the FCS files.

`--transform [<parameters>]` Transform parameters using the default biexponential transformation. Parameters are passed in as a comma separated list, e.g. `--transform FL1-A,FL2-A,FL3-A`

`--rscript=<rscript>` if additional processing needs to be done, pass in an R script. The script can work with the GatingSet variable `gs`. It is run after `compensate` and `transform`.

`--gs=<gatingset>` optionally read from a GatingSet rather than FCS files. Can be combined with `--rscript`. GatingSet data are already compensated and transformed.




