---
title: "Using OpenCyto Command Line Tools"
author: "Greg Finak"
date: "`r Sys.Date()`"
output: 
   rmarkdown::html_vignette:
      toc: true
vignette: >
  %\VignetteIndexEntry{Using OpenCytoCLI}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
# OpenCyto command line tools
A thin wrapper around the opencyto  framework leveraging littler.

Run simple flow cytometery processing tasks using opencyto from the command line.

## Notes

Currently openCyto command line tools work with Linux and Mac OS X

OpenCyto command line tools expects well standardized and annotated data free of typos in keywords, markers and channels.

## Errors? Bugs? Feature Requrests?

Open an issue, provide a reproducible example.

## Install OpenCyto Command Line Tools

To install the package run the following in R:


```{r eval=FALSE,echo=TRUE}
library(devtools)
install_github("RGLab/opencytoCL")
```

You will need all the dependencies:

- openCyto
- optparse
- littler

## Linux and MacOS X

Then, in R:

```{r, eval=FALSE,results="asis"}
install_openCytoCL()
```

## Usage
For help, at the command prompt, type:

```{bash, comment = "",size="small",tidy=TRUE}
opencyto -h
```

### parse
Import a FlowJo workspace and related FCS files and create a GatingSet.

#### options

`<workspace>` the location of the workspace

`<fcs_read_from>` the location of the fcs files

`<save_to>` the name of the gating set to save

`--group=<group_number>` option to specify the group number to import. Print all group names using `load` (see below).

`--annotate=<keywords>` optionally annotate by importing keywords. Keywords should be passed in as a quoted, comma separated list. If parsing FCS files, the keywords are taken from the FCS files.  If parsing a workspace, the keywords are taken from the workspace. Spaces in keywords should be replaced with `:`.

`--stats` Extract population statistics and write them to a file.

### load
Load a FlowJo workspace and print the sample groups in therein.

#### options

`--gs=<gatingset>` the name of the gating set to load. Created using `parse` (see above) or `process` (see below) 

### gate
Gate a GatingSet using a template, saving the results. Optionally extract population statistics. Any existing gates in the GatingSet are removed. 

#### options

`--gs=<gatingset>`  A GatingSet on which to perform automated gating.

`--template=<template>`  The name of an openCyto template csv file.

`--stats`  Extract population statistics and write them to a file.

### process

`<fcs_read_from>` location of files to process

`<save_to>` name of gating set to save to

`--compensate` perform compensation using the matrix in the FCS files.

`--transform [<parameters>]` Transform parameters using the default biexponential transformation. Parameters are passed in as a comma separated list, e.g. `--transform FL1-A,FL2-A,FL3-A`

`--rscript=<rscript>` if additional processing needs to be done, pass in an R script. The script can work with the GatingSet variable `gs`. It is run after `compensate` and `transform`.

`--gs=<gatingset>` optionally read from a GatingSet rather than FCS files. Can be combined with `--rscript`. GatingSet data are already compensated and transformed.

# Examples

Install `flowWorkspaceData` from BioConductor.

```{r results="hide"}
if(!library(flowWorkspaceData,logical.return = TRUE)){
		source("https://bioconductor.org/biocLite.R")
		biocLite("flowWorkspaceData")
}
```

## Create a gating set from FCS files.

```{r echo=FALSE,eval=TRUE,results="hide"}
fcs_path = system.file(package="flowWorkspaceData","extdata")
tmp = tempdir()
try(file.remove(list.files("/tmp/cytotrol",full=TRUE)))
try(file.remove("/tmp/cytotrol"))
file.copy(list.files(path=fcs_path,pattern="CytoTrol.*",full=TRUE),to = tmp)
```

```{r, echo=FALSE, results='hide'}
system(command=paste0("opencyto process ",tmp,"  /tmp/cytotrol --compensate --transform B710-A,R660-A,R780-A,V450-A,V545-A,G560-A,G780-A --annotate='$FIL'"))
```

Next, run command line opencyto to process FCS files into a compensated and transformed gatings set.

```{sh eval=FALSE,echo=TRUE,results="hide"}
opencyto process /Library/Frameworks/R.framework/Versions/3.4/Resources/library/flowWorkspaceData/extdata /tmp/cytotrol_gatingset --compensate --transform --annotate="$FIL"
```

```{sh eval=FALSE}
Loading opencyto.
All FCS files have the same following channels:
FSC-A
FSC-H
FSC-W
SSC-A
B710-A
R660-A
R780-A
V450-A
V545-A
G560-A
G780-A
Time
write CytoTrol_CytoTrol_1.fcs to empty cdf slot...
write CytoTrol_CytoTrol_2.fcs to empty cdf slot...
done!
write CytoTrol_CytoTrol_1.fcs to empty cdf slot...
write CytoTrol_CytoTrol_2.fcs to empty cdf slot...
..done!
Transforming B710-A R660-A R780-A V450-A V545-A G560-A G780-A
saving ncdf...
saving tree object...
saving R object...
Done
To reload it, use 'load_gs' function

Exiting.
```

Our gating set now lives at `/tmp/cytotrol`. 

## Create a gating set from a FlowJo workspace file.

```{r}
manual_xml = list.files(fcs_path,pattern="manual.xml",full=TRUE)
```

The FlowJo xml file describing the manually gated data in `flowWorskpaceData` is at `r manual_xml`. We can parse it and extract cell population statistics.

```{r,echo=FALSE,results="hide"}
system(paste0("opencyto parse ",manual_xml))
```

```{r echo=FALSE,eval=TRUE,results="hide",error=FALSE,warning=FALSE}
try(file.remove(list.files("/tmp/cytotrol_manual",full=TRUE)))
try(file.remove("/tmp/cytotrol_manual"),silent=TRUE)
```

```{sh eval=FALSE}
Loading opencyto.

Found sample groups
	 1. All Samples 
	 2. B-cell 
	 3. DC 
	 4. T-cell 
	 5. Thelper 
	 6. Treg 
Parse using:
opencyto parse /Library/Frameworks/R.framework/Versions/3.4/Resources/library/flowWorkspaceData/extdata/manual.xml <read_from> <save_to> --group=<groupnumber>
Workspace not parsed.
```

We see a number of sample groups. We'll load the T-cell group (4).

```{r echo=FALSE,eval=TRUE,results="hide",error=FALSE,warning=FALSE}
try(file.remove(list.files("/tmp/cytotrol_manual",full=TRUE)))
try(file.remove("/tmp/cytotrol_manual"),silent=TRUE)
```

```{r echo=FALSE,results="hide"}
file.remove(files = list.files(path = getwd(),pattern = "Rtmp.*\\.csv",full.names=TRUE))
```


```{r echo=TRUE,results="hide"}
system(command=paste0("opencyto parse ",manual_xml," ",fcs_path,"  /tmp/cytotrol_manual --stats --group=4"))
```

```{sh ,eval=FALSE}
Loading opencyto.

Found sample groups
	 1. All Samples 
	 2. B-cell 
	 3. DC 
	 4. T-cell 
	 5. Thelper 
	 6. Treg 
Parsing group 4 from
/Library/Frameworks/R.framework/Versions/3.4/Resources/library/flowWorkspaceData/extdata
Parsing 2 samples
mac version of flowJo workspace recognized.
Creating ncdfFlowSet...
All FCS files have the same following channels:
FSC-A
FSC-H
FSC-W
SSC-A
B710-A
R660-A
R780-A
V450-A
V545-A
G560-A
G780-A
Time
done!
loading data: /Library/Frameworks/R.framework/Versions/3.4/Resources/library/flowWorkspaceData/extdata/CytoTrol_CytoTrol_1.fcs
Compensating
gating ...
write CytoTrol_CytoTrol_1.fcs_119531 to empty cdf slot...
loading data: /Library/Frameworks/R.framework/Versions/3.4/Resources/library/flowWorkspaceData/extdata/CytoTrol_CytoTrol_2.fcs
Compensating
gating ...
write CytoTrol_CytoTrol_2.fcs_115728 to empty cdf slot...
done!
saving ncdf...
saving tree object...
saving R object...
Done
To reload it, use 'load_gs' function

Done!
Exiting.
```

The cell population statistics are stored in the local directory:

```{r echo=FALSE}
getwd()
```

```{r echo=FALSE}
files = list.files(path = getwd(),pattern = "Rtmp.*\\.csv",full.names=TRUE)
if(length(files)==1){
	stats = data.table::fread(input = files)
	head(stats)
}
```


```{r echo=FALSE,results="hide"}
#clean up afterwards
file.remove(files = list.files(path = getwd(),pattern = "Rtmp.*\\.csv",full.names=TRUE))
```

